---
title: "Homework 3 solutions"
author: "Matthew Angulo"
date: "October 10, 2020"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)


theme_set(theme_minimal()+theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colur = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```


## Problem 1

```{r}
data("instacart") 
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns 

Observations are at the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles, and which are most items from?

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```


Let's make a plot 

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n >10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n))+
  geom_point()+
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust=1))
```

Let's make a table!

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, desc(n)) %>% 
  knitr::kable()
```


Apples vs ice cream

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
  

```


## Question 2

```{r part 2.1}
accel_df = read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    day = factor(day, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
  )) %>% 
  arrange(day, week) %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "activity_count"
  ) %>% 
  mutate(
    weekday_weekend = case_when(
      day %in% c("Saturday", "Sunday") ~ "weekend",
      day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~ "weekday",
      TRUE ~ "")
  ) %>% 
  mutate(
     weekday_weekend = factor(weekday_weekend),
     minute = as.numeric(minute)
  ) %>% 
  relocate(week, day_id, day, weekday_weekend)

accel_df
```

This dataset contains the amount of activity one subject exerted over a number of days. The accel_df has `r ncol(accel_df)` variables that include day, whether it was the weekend or not, the minute of the day, and the amount of activity recorded. In this dataset, there were`r nrow(accel_df)` observations - one for each second of the day for 35 days.   


```{r part 2.2}
accel_df %>% 
  group_by(week, day) %>% 
  summarize(total = sum(activity_count)) %>% 
  knitr::kable()
```

Analyzing the total activity count for the 35 days shows that there are several outliers. On two Saturdays, the subject recorded the baseline amount of activity. The first Monday of the trial also recorded a very low amount of activity. 



```{r part 2.3}
accel_df %>% 
  ggplot(aes(x = minute, y =activity_count, color = day))+
  geom_line(alpha = .7)
```

The subject recorded little activity between 10:00AM (or ~600 minutes) and 6:30PM (or ~1100 minutes). The subject did have a frequent high peak during the work week around 9pm. Sundays the subject exerted more activity during the middle fo the day in comparison to other days. 



## Question 3

```{r}
library(p8105.datasets)
library(hexbin)
library(ggridges)
library(patchwork)
data("ny_noaa")
```


```{r part 3.1}
ny_noaa_df = 
  ny_noaa %>% 
  janitor::clean_names() %>%
  separate(date, into = c("year", "month", "day"), convert = TRUE) %>% 
  mutate(
    tmin = as.numeric(tmin),
    tmax = as.numeric(tmax),
    prcp = prcp * 0.1,
    tmax = tmax * 0.1,
    tmin = tmin * 0.1
    )

count(ny_noaa_df, snow) %>% 
  arrange(-n)
  
```

This is a large dataset that has `r nrow(ny_noaa_df)` observations and `r ncol(ny_noaa_df)` variables including: weather station, day, month, year, min and max temperature, amount of rainfall, amount of snowfall, and snow density. There are numerous stations that are missing records on precipitation, temperature, and snowfall throughout the years. Precipation values were converted to mm from tenths of

The most common amount of reported snowfall was either 0mm or NA. Certainly the expected distribution since the NY regions experiences more days without snowfall than days with it. 




```{r part 3.2}
panel_df= 
  ny_noaa_df %>% 
  filter(month %in% c("1","7")) %>% 
  group_by(year, month, id) %>% 
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = mean_tmax, group = id, color = id))+
  geom_point(alpha = .5)+
  geom_path()+
  facet_grid(. ~ month)+
  labs(title = "Mean max temperature for January and July across stations and years", x = "year", y = "average maximum temperature (C)")+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

panel_df
```

The mean maximum temperature among these weather stations shows that January 1994 and 2004 were particularly cold winters. However, the max temperatures in January at these weather stations seems to be trending warmer. 

The July 2010 max temperatures were some of the highest recorded for that month since this dataset first started recording.



```{r part 3.3}
hex_panel_df = 
  ny_noaa_df %>% 
  ggplot(aes(x = tmin, y = tmax))+
  geom_hex()+
  labs(title = "Daily max temperatures versus daily min temperatures")+
  guides(fill = guide_colorbar(barwidth = 12))+
  theme(plot.title = element_text(hjust = 0.5))

hex_panel_df


box_panel_df =
  ny_noaa_df %>% 
  filter(snow %in% c(1:99)) %>% 
  ggplot(aes( x = snow, y = year, group = year))+
  geom_boxplot()+
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
  labs(title = "Yearly amount of snowfall")

box_panel_df

hex_panel_df + box_panel_df

```

